The how-to I always wanted
========================================================
For this assignment only some columns of a .csv table are interesting. Those columns will be extracted first. Afterwards there will be two new columns added to the dataframe. For the first new column the timedifference between two measurements (two rows) will be calculated. the other new column will contain the unit of the time difference. 

First of all a general function to insert new columns is implemented: 

```{r}

insert.col <- function(input.data.frame, col.num, col.content) {
      input.data.frame[(col.num+1):(ncol(input.data.frame)+1), ] <- 
        input.data.frame[col.num:ncol(input.data.frame),]
    input.data.frame[col.num, ] <- col.content
    return(input.data.frame)
    }
```

The following function converts a string to a proper date/time information. 
```{r}
convert.date <- function(row.num)
  row.date <- as.POSIXct(strptime(input.data.frame[row.num, 1], "%Y-%m-%d %H:%M:%S"))
```

The original data is put into a new dataframe and the interesting columns are extracted

```{r}
setwd ("C:/Users/Alice/Documents/__Studium/Master_Marburg/1.Semester/Data_Management/DM09_Markdownlanguage")
data.complete <- read.table ("csv_example_in.csv", header=TRUE, sep=";", quote="\"'", dec=".", na.strings = "nan", stringsAsFactors=FALSE)
input.data.frame <- data.complete[,c(1,2,4,9)]
```
<-- Errormessage: attempt to use zero-length variable name


Two columns are added into the dataframe: 
```{r}
col.num <- 2
for (i in (nrow(input.data.frame)-1){
  date1 <- convert.date(i)
  date2 <- convert.date(i+1)
  time.dif <- date2-date1
  col.content <- c(col.content, time.dif)}

  work1.data.frame <- insert.col(data.complete[,c(1:2,4,9)], 2, col.content) # kann statt data complete[,c(1:2,4,9)] auch einfach input.data.frame stehen?
```

```{r}
col.num <- 3
col.content <- # unit
  work2.data.frame <- insert.col(work1.data.frame, 3, col.content)
```

The new dataframe is written into a .csv table
```{r}
write.csv(work2.data.frame, file= "csv_example_out")
```
