W09-1: Data manipulation
========================================================
In this exercise, a climate station  file is read an certain columns are
selected for writing to an output. In addition, a new column is added giving
the time difference between the individual measurements.

We begin this task with reading the csv file and selecting the appropriate
data subset from the input. As one can see in the output below, the first
column consists of strings which represent data/time values.
```{r}
path <- "D:/active/moc/dm/programs/github/meg-dm-2013-winter/tnauss/data/"

# Read csv file to data frame
my.data <- read.table(paste(path,"csv_example_in.csv", sep=""), 
                      header=TRUE, sep=",", quote="\"'", dec=".",
                      na.strings = "nan", stringsAsFactors=FALSE)

my.subset <- my.data[,c(1:2,4,9)]
print(my.subset)
```
To compute the time difference, one has to convert the data/time strings to
data/time values. This is realized by the following code.
```{r}
my.subset[,1] <- as.POSIXct(strptime(my.subset[,1], "%Y-%m-%d %H:%M:%S"))
```
Finally, we have to insert the time difference between each time step as second
column. As third column, the unit of the time difference should be added.

The column insertion is implemented using a function which first copies the
content of every column equal or larger than the insert position to the column
next to it on the right side. Afterwards the new column is inserted at the
appropriate position (overwriting the content of this column but that is ok
since we moved the content to the right). After moving the columns we have
to take care of the column names. They are moved to the right/overwritten in
the same manner as the content of the columns.

To include the time difference in the output data frame, we first compute the
differences with the diff function. Since the difference can not be computed 
for the last item, the vector has to be extended by one number (0 in this case) 
to fit the dimension of the data frame. We do that in two subsequent rows to 
still have the original content in the time.differnce.

After computing the time difference, we can insert the difference using the 
pre-defined function. 
```{r}
col.insert <- function(input.data.frame, col.pos, col.content, col.name) {
  input.data.frame[,(col.pos+1):(ncol(input.data.frame)+1)] <-
    input.data.frame[,col.pos:ncol(input.data.frame)]
  input.data.frame[,col.pos] <- col.content
  colnames(input.data.frame)[(col.pos+1):ncol(input.data.frame)] <-
    colnames(input.data.frame)[col.pos:(ncol(input.data.frame)-1)]
  colnames(input.data.frame)[col.pos] <- col.name
  return(input.data.frame)
}

time.difference <- diff(my.subset[,1])
col.content <- c(time.difference, 0)
output.data.frame <- col.insert(my.subset, 2, col.content, "Timedifference")
```
Almost finished. We just need to insert the  unit of the
time differnce in a third column. Since we stored the time diffrence in a 
variable called time.difference. Let's have a look at its structure:
```{r}
str(time.difference)
```
The variable has an attribute "units" which gives us what we need. We can get
the attribute value using the attr function. Once we have it, we just use our
self-defined function again to insert the column.
```{r}
attr(time.difference, "units")
output.data.frame <- col.insert(output.data.frame, 3, attr(test, "units"), "TDUnit")
print(output.data.frame)
```
Ok, let's just write the stuff to a csv file and call it a day.
```{r}
write.table(output.data.frame, paste(path,"output_with_new_cols.csv", sep=""), 
            sep=",", dec=".", quote=FALSE, row.names=FALSE)
```
